Individualized Measures of Partisan Fairness and Utility in Redistricting (??)

### 2021-08-31 paper/paper.Rmd
cleanup + fig labels

Votes may be tallied at the district level, either in total or by party: \begin{align*}
    \votes&: \Distrs\times\Elecs \to \N, \qquad 
    \votes(d, v) = \sum_{i\in q^{-1}(d)} \ind_P(v(i)) \\
    \votes^{(p)}&: \Distrs\times\Elecs\times\Plans \to \N, \qquad 
    \votes^{(p)}(d, v, q) = \sum_{i\in q^{-1}(d)} \ind_{\{p\}}(v(i)).
\end{align*} The political party of the candidate with the most votes wins each district, as captured by the function $$
    \win: \Distrs\times\Elecs\times\Plans\to P, \qquad 
    \win(d, v, q) = \argmax_{p\in\Parties} \votes^{(p)}(d, v, q).
$$Thus an individual voter $i\in\Voters$ is represented after election $v$ by political party $\win(q(i), v, q)$, and each political party can be said to win a number of seats, formalized as$$
    \seats:\Parties\times\Elecs\times\Plans\to\N, 
    \qquad \seats(p,v,q) = \sum_{d\in\Distrs} \ind_{\{p\}}(\win(d,v,q)).
$$

@@

What kind of electoral systems does this utility function encourage, if the goal is maximizing total voter utility (or, equivalently, average voter utility)?
It depends on the form of $h$, as the following result describes.

@@

\end{restatable}
```
Thus, if we are willing to adopt the logarithm as the functional specification of $u^{(G)}$, then this global outcome utility measure will encourage districting systems that create proportionality between party vote shares and seat shares.

@@

The expected harm to a specific voter then becomes $$
   \E[H\mid I=i] = \Pr\qty(V(i)\neq\win(Q(i), V, Q)
    \text{ and }V(i)=\win(\tilde Q(i), V, \tilde Q)),
$$

@@

In practice, few voters will experience zero harm.
Most voters living in a district with a political preference not matching their own could draw a districting plan, however contrived, which puts them in a district dominated by voters of their party.
So merely targeting zero harm for all voters is not an adequate representation of fairness.
Rather, intuitively, a districting plan is fair if it does not harm the voters of one party more than another, on average.
**[CM: need to rephrase here. awkward]**

@@

Continuing with the above notation, the average fairness of $Q$ versus counterfactual $\tilde Q$ may be defined as $$
    F = \E[H\mid V(I) = A] - \E[H\mid V(I) = B];
$$obviously the ordering of $A$ and $B$ is immaterial.
Informally, this quantity is the difference in the probability that a voter of each party is represented by the other party but counterfactually would have been represented by their own party.

>>>>

Votes may be tallied at the district level, by party, by defining for each party $p\in\Parties$ $$
    \votes^{(p)}: \Distrs\times\Elecs\times\Plans \to \N, \qquad 
    \votes^{(p)}(d, v, q) = \sum_{i\in q^{-1}(d)} \ind_{\{p\}}(v(i)).
$$ The political party of the candidate with the most votes wins each district, as captured by the function $$
    \win: \Distrs\times\Elecs\times\Plans\to P, \qquad 
    \win(d, v, q) = \argmax_{p\in\Parties} \votes^{(p)}(d, v, q).
$$Thus an individual voter $i\in\Voters$ is represented after election $v$ by political party $\win(q(i), v, q)$, and each political party can be said to win a number of seats, formalized as$$
    \seats:\Parties\times\Elecs\times\Plans\to\N, 
    \qquad \seats(p,v,q) = \sum_{d\in\Distrs} \ind_{\{p\}}(\win(d,v,q)).
$$

@@

What kind of electoral systems does this utility function encourage, if the goal is maximizing total voter utility (or, equivalently, average voter utility)?
It depends on the form of $g$, as the following result describes.

@@

\end{restatable}
```
Thus, if we are willing to adopt the logarithm for the functional specification of $u^{(G)}$, then this global outcome utility measure will encourage districting systems that create proportionality between party vote shares and seat shares.

@@

The expected harm to an individual $i\in\Voters$ voter then becomes $$
   \E[H\mid I=i] = \Pr\qty(V(i)\neq\win(Q(i), V, Q)
    \text{ and }V(i)=\win(\tilde Q(i), V, \tilde Q)),
$$

@@

@@

Intuitively, a districting plan is fair if it does not harm the voters of one party more than another, on average.
This is easily expressed using our notation above; the average fairness of $Q$ versus counterfactual $\tilde Q$ may be defined as $$
    F = \E[H\mid V(I) = A] - \E[H\mid V(I) = B];
$$obviously the ordering of $A$ and $B$ is immaterial.
Informally, this quantity is the difference in the probability that a voter of each party is represented by the other party but counterfactually would have been represented by their own party.
